### 2022-09-28 main/chapter_context.tex
fix higgs H mixup; write on naturalness

Other mysteries of cosmology include:
the positive cosmological constant,
the inflation of the early universe,
and baryogenesis~\cite{
wells2020discovery,
riess1998observational
}.
Answers to these mysteries all lie somewhere between uncertain and unknown,
but may not be entirely out of reach;
for example, the Higgs field could plausibly have driven
inflation~\cite{bezrukov2008higgs},
and the Standard Model does slightly violate $C$ and $CP$, both of which
are required for baryogenesis~\cite{sakharov1991re}.

@@

Certain Standard Model structures are quite general and parametrized such
that parts can be tuned to describe nature.
For example, the strong interaction includes a parameter $\theta_\mathrm{CP}$ whose
non-zero values would induce $\mathrm{CP}$ violation.
Observations find it very close to zero, suggesting that nature does not use
that freedom, so one promising feature of axion-like models is their
implication that $\theta_\mathrm{CP} = 0$~\cite{
kim2008axions,
thomson2013modern,
martin2017particle
}.

@@

\paragraph{Aesthetics.}
\TODO{naturalness and fine-tuning~\cite{wells2020discovery}}
\TODO{strong CP problem}
\TODO{hierarchy problem}

Naturalness is opinion and only nature decides what is really natural.

@@

% using https://www.tablesgenerator.com/latex_tables
\begin{table}[tp]
\centering
\begin{tabular}{rrc}
\multirow{5}{*}{bosons}   & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
                          &                               & \makebox[5em]{\hfill $\tilde d$ \hfill $\tilde s$ \hfill $\tilde b$ \hfill}                  \\[1ex]
                          & \multirow{2}{*}{sleptions}    & \makebox[5em]{\hfill $\tilde e$ \hfill $\tilde \mu$ \hfill $\tilde \tau$ \hfill}             \\
                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[1ex]
                          & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[3ex]
\multirow{3}{*}{fermions} & strong                        & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
                          & \multirow{2}{*}{electroweak}  & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
                          &                               & \makebox[7em]{\hfill $\tilde N_1\!$ \hfill $\tilde N_2\!$ \hfill$\tilde N_3\!$ \hfill$\tilde N_4\!$ \hfill} \\
\end{tabular}
\caption[Particles of the Minimal Supersymmetric Standard Model]{%
Particles of the Minimal Supersymmetric Standard Model
\TODO{check and cite}
\TODO{relate to Standard Model}
\TODO{adds goldstino and gravitino}
\TODO{alternative $\chargino_i = \tilde C^\pm_i$, $\neutralino_i = N_i$}
\TODO{In this table, all bosons are spin-$0$ scalars and all fermions are spin-$1/2$.}
\TODO{The bracketed $h$ corresponds to the Standard Model Higgs boson.}
}
\label{tab:context_particles_mssm}
\end{table}

@@

The Higgs boson couples to an extraordinary variety of fundamental particles.
For its discovery alone, \atlas's designers achieved sensitivity across the
entire $100$ to $1000\,\eV[G]$ window by targetting decays including
$H\to b\bar b$,
$H\to \gamma\gamma$,
$H\to ZZ^* \to 4\ell$, and
$H\to ZZ \to \ell\ell\nu\nu$;
these channels demand precise measurements of leptons, photons, and
$b$-jets (jets seeded from $b$ quarks), as well as sensitivity to the missing
transverse momentum of invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays
at the very highest masses.

Both \atlas\ and \cms\ have now found a Higgs boson candidate at
$m(H) = 125\,\eV[G]$~\cite{
atlas2012higgs,
atlas2012combined,
cms2012higgs
}
and confirmed all of its tested Standard Model properties~\cite{
combined2016higgs,
atlas2022ten,
cms2022ten
}
such as its production mechanisms, decay rates,
spin, and parity~\cite{
HIGG-2013-01,
HIGG-2013-17,
HIGG-2014-06
}.

@@

Indeed, the main search presented in this thesis~\cite{atlas2022searches}
joins nicely with the pre-existing needs for sensitivity to signatures
including
$H \to b\bar b$
and
$ZZ \to \ell\ell\nu\nu$.

>>>>

Other mysteries of cosmology include:
the small positive cosmological constant,
the inflation of the early universe,
and baryogenesis~\cite{
wells2020discovery,
riess1998observational
}.
Answers to these mysteries all lie somewhere between uncertain and unknown,
but may not be entirely out of reach;
for example, the Higgs field could plausibly have driven
inflation~\cite{bezrukov2008higgs},
and the Standard Model does slightly violate $C$ and $CP$, both of which
are required for baryogenesis~\cite{sakharov1991re}.

@@

Certain Standard Model structures are quite general and parametrized such
that parts can be tuned to describe nature.
For example, the strong interaction includes a parameter $\theta_\mathrm{CP}$
whose non-zero values would induce $\mathrm{CP}$ violation in the strong
sector.
Observations show that it cannot be far from zero, and one promising feature of
axion-like models is their implication that $\theta_\mathrm{CP} = 0$~\cite{
kim2008axions,
thomson2013modern,
martin2017particle
}.

@@

\paragraph{Aesthetics.}
In the search for improved models, heuristic criteria can help to identify
appealing candidates.
Two such criteria are `fine-tuning' and `naturalness', both of which are
aesthetic judgements about dimensionless model parameters.
Naturalness is opinion; only nature decides what is really natural,
but these criteria can still guide the discourse of our scientific searches
towards useful directions~\cite{wells2020discovery}.

Fine-tuning criteria dislike tight constraints, and naturalness contains
various ideas of which one is that nice numbers should have moderate orders of
magnitude~\cite{giudice2008naturally}.
For example, $\theta_\mathrm{CP}$ is unattractively fine-tuned within its
experimental constraint of $\theta_\mathrm{CP} \lesssim 10^{-11}$, and that
large $11$ in its exponent defies naturalness.

Another example is the hierarchy problem of the Higgs boson mass $m(h)$,
which notes that the energy scale of new heavy particles linearly modifies the
$m(h)$ through direct or indirect couplings.
With any particles near the Planck scale $\sim 10^{18}\,\eV[G]$, implementing
the observed $m(h) \sim 10^{2}\,\eV[G]$ therefore requires finely-tuned
cancellations between positive and negative additions.
New fermions subtract from $m(h)$ and new scalars add by similar amounts,
so new models can implement this fine-tuning by inserting only matched pairs of
fermions and scalars --- this is the strategy of supersymmetric models.

@@

% made using:
% https://www.tablesgenerator.com/latex_tables
\begin{table}[tp]
\centering
\begin{tabular}{rrc}
\multirow{5}{*}{bosons}   & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
                          &                               & \makebox[5em]{\hfill $\tilde d$ \hfill $\tilde s$ \hfill $\tilde b$ \hfill}                  \\[1ex]
                          & \multirow{2}{*}{sleptions}    & \makebox[5em]{\hfill $\tilde e$ \hfill $\tilde \mu$ \hfill $\tilde \tau$ \hfill}             \\
                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[1ex]
                          & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[3ex]
\multirow{3}{*}{fermions} & strong                        & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
                          & \multirow{2}{*}{electroweak}  & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
                          &                               & \makebox[7em]{\hfill $\tilde N_1\!$ \hfill $\tilde N_2\!$ \hfill$\tilde N_3\!$ \hfill$\tilde N_4\!$ \hfill} \\
\end{tabular}
\caption[Particles of the Minimal Supersymmetric Standard Model]{%
Particles of the Minimal Supersymmetric Standard Model
\TODO{check and cite}
\TODO{relate to Standard Model}
\TODO{adds goldstino and gravitino}
\TODO{alternative $\chargino_i = \tilde C^\pm_i$, $\neutralino_i = N_i$}
\TODO{In this table, all bosons are spin-$0$ scalars and all fermions are spin-$1/2$.}
\TODO{The bracketed $h$ corresponds to the Standard Model Higgs boson.}
}
\label{tab:context_particles_mssm}
\end{table}

@@

The Higgs boson couples to an extraordinary variety of fundamental particles.
For its discovery alone, \atlas's designers achieved sensitivity across the
entire $100$ to $1000\,\eV[G]$ window by targetting decays including
$h \to b\bar b$,
$h \to \gamma\gamma$,
$h \to ZZ^* \to 4\ell$, and
$h \to ZZ \to \ell\ell\nu\nu$;
these channels demand precise measurements of leptons, photons, and
$b$-jets (jets seeded from $b$ quarks), as well as sensitivity to the missing
transverse momentum of invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays
at the very highest masses.

Both \atlas\ and \cms\ have now found a Higgs boson candidate at
$m(h) = 125\,\eV[G]$~\cite{
atlas2012higgs,
atlas2012combined,
cms2012higgs
}
and confirmed all of its tested Standard Model properties~\cite{
combined2016higgs,
atlas2022ten,
cms2022ten
}
such as its production mechanisms, decay rates,
spin, and parity~\cite{
HIGG-2013-01,
HIGG-2013-17,
HIGG-2014-06
}.

@@

Indeed, the main search presented in this thesis~\cite{atlas2022searches}
joins nicely with the pre-existing needs for sensitivity to signatures
including
$h \to b\bar b$
and
$ZZ \to \ell\ell\nu\nu$.
